\Large{Adversarial Variational Optimization of Non-Differentiable Simulators}

### 2017-06-27 tex/paper.tex
save

As illustrative experiments, we evaluate adversarial variational
optimization on well-known distributions for continuous and discrete data.
We consider these distributions as parameterized simulators, from which
we can only generate data.

For continuous data, we first consider inference for
a univariate Gaussian with unknown mean and
unknown variance. The observed data is sampled from a univariate Gaussian with
mean $\mu = 1$ and variance $\sigma^2 = 0.5^2$. Algorithm~\ref{alg:avo} is run
for 300 epochs with mini-batches of size $M=64$ and the following configuration.
For the critic, we use a 3-layer MLP with 10 hidden nodes per layer and ReLU
activations. At each epoch, Adam is run for $n_\text{critic}=100$ iterations
with a step size $\alpha=0.01$, decay rates $\beta_1=\beta_2=0.5$ and its inner
first and second moment vectors reset at each epoch in order to avoid building
momentum in staled directions.  For modeling $\mu$ and $\sigma^2$, we use
univariate Gaussian proposal distributions $q(\mu|\psi)$ and $q(\log
\sigma|\psi)$ initialized with zero mean and unit variance. At each epoch,
parameters $\psi$ are updated by making one Adam step, with $\alpha=0.01$ and
$\beta_1=\beta_2=0.1$. Penalty coefficients are set to $\lambda=0.025$ and
$\gamma=5$.

\begin{figure}
    \centering
    \vspace{-0.8cm}
    \includegraphics[width=0.48\textwidth]{figures/gaussian-gamma=5.pdf}
    \vspace{-1cm}
    \caption{Toy problem on continuous data: Gaussian with unknown mean and unknown variance.}\label{fig:gaussian}
\end{figure}
\begin{figure}
    \centering
    \vspace{-0.8cm}
    \includegraphics[width=0.48\textwidth]{figures/poisson-gamma=5.pdf}
    \vspace{-1cm}
    \caption{Toy problem on discrete data: Poisson with unknown scale.}\label{fig:poisson}
\end{figure}

The upper plot of Figure~\ref{fig:gaussian} illustrates the resulting model
after the optimization procedure. As expected from adversarial training, the
model density $p(\mathbf{x}|\psi)$ closely matches the true density
$p_r(\mathbf{x})$, as shown by the aligned histograms in blue and orange. Under
the effect of the entropy penalty $\gamma H(q(\theta|\psi))$, we also observe
that the proposal distributions have correctly concentrated their mass around
$\mu=1$ and $\log \sigma=-0.69$. The bottom plot of Figure~\ref{fig:gaussian}
shows an empirical estimate of $-U_d$ with respect to the epoch. We observe
that it quickly falls towards $0$, which indicates that
$\mathbb{E}_{\tilde{\mathbf{x}} \sim p(\mathbf{x}|\theta)} [d(\tilde{\mathbf{x}};\phi)] \approx \mathbb{E}_{\mathbf{x} \sim p_r(\mathbf{x})} [d(\mathbf{x};\phi)]$
and that the critic cannot distinguish between true and model data.

>>>>

As illustrative experiments, we evaluate adversarial variational optimization on
well-known distributions for continuous and discrete data. We artificially
consider these distributions as parameterized black box simulators, from which
we can only generate data.

For continuous data, we first consider inference for
a univariate Gaussian with unknown mean and
unknown variance. The observed data is sampled from a univariate Gaussian with
mean $\mu = 1$ and variance $\sigma^2 = 0.5^2$. Algorithm~\ref{alg:avo} is run
for 300 epochs with mini-batches of size $M=64$ and the following configuration.
For the critic, we use a 3-layer MLP with 10 hidden nodes per layer and ReLU
activations. At each epoch, Adam is run for $n_\text{critic}=100$ iterations
with a step size $\alpha=0.01$, decay rates $\beta_1=\beta_2=0.5$ and its inner
first and second moment vectors reset at each epoch in order to avoid building
momentum in staled directions.  For modeling $\mu$ and $\sigma^2$, we use
factored univariate Gaussian proposal distributions $q(\mu|\psi)$ and $q(\log
\sigma|\psi)$ initialized with zero mean and unit variance. At each epoch,
parameters $\psi$ are updated by making one Adam step, with $\alpha=0.01$ and
$\beta_1=\beta_2=0.1$. Penalty coefficients are set to $\lambda=0.025$ and
$\gamma=5$.

\begin{figure}
    \centering
    \vspace{-0.8cm}
    \includegraphics[width=0.48\textwidth]{figures/gaussian-gamma=5.pdf}
    \vspace{-1cm}
    \caption{Toy problem on continuous data: Gaussian with unknown mean and unknown variance.}\label{fig:gaussian}
\end{figure}
\begin{figure}
    \centering
    \vspace{-0.5cm}
    \includegraphics[width=0.48\textwidth]{figures/poisson-gamma=5.pdf}
    \vspace{-1cm}
    \caption{Toy problem on discrete data: Poisson with unknown scale.}\label{fig:poisson}
\end{figure}

The upper plot of Figure~\ref{fig:gaussian} illustrates the resulting model
after the optimization procedure. As theoretically expected from adversarial training, the
model density $p(\mathbf{x}|\psi)$ closely matches the true density
$p_r(\mathbf{x})$, as shown by the aligned histograms in blue and orange. Under
the effect of the entropy penalty $\gamma H(q(\theta|\psi))$, we also observe
that the proposal distributions have correctly concentrated their mass around
$\mu=1$ and $\log \sigma=-0.69$. The bottom plot of Figure~\ref{fig:gaussian}
shows an empirical estimate of $-U_d$ with respect to the epoch. We observe
that it quickly falls towards $0$, which indicates that
$\mathbb{E}_{\tilde{\mathbf{x}} \sim p(\mathbf{x}|\theta)} [d(\tilde{\mathbf{x}};\phi)] \approx \mathbb{E}_{\mathbf{x} \sim p_r(\mathbf{x})} [d(\mathbf{x};\phi)]$
and that the critic cannot distinguish between true and model data.
